---
output: github_document
---

<!-- README.md is generated from README.Rmd. Edit that file, then knit. -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  eval = FALSE
)
```

## virustotal: R Client for the VirusTotal API v3

[![R-CMD-check](https://github.com/themains/virustotal/actions/workflows/R-CMD-check.yml/badge.svg)](https://github.com/themains/virustotal/actions/workflows/R-CMD-check.yml)
[![lint](https://github.com/themains/virustotal/actions/workflows/lint.yml/badge.svg)](https://github.com/themains/virustotal/actions/workflows/lint.yml)
[![codecov](https://codecov.io/gh/themains/virustotal/branch/master/graph/badge.svg)](https://app.codecov.io/gh/themains/virustotal)
[![CRAN status](https://www.r-pkg.org/badges/version/virustotal)](https://cran.r-project.org/package=virustotal)
[![lifecycle](https://img.shields.io/badge/lifecycle-stable-brightgreen.svg)](https://lifecycle.r-lib.org/articles/stages.html#stable)
![](https://cranlogs.r-pkg.org/badges/grand-total/virustotal)

Use [VirusTotal](https://www.virustotal.com), a Google service that analyzes
files and URLs for viruses, worms, trojans and other malware, categorizes the
content hosted by a domain using a variety of prominent services, and provides
passive DNS information, among other things.

The package covers the VirusTotal API v3: file, URL, domain and IP reports;
scanning and rescanning; IoC relationships; sandbox behaviour artifacts;
comments and votes. Requests are paced to the public API's 4-per-minute
allowance automatically, transient failures are retried honoring the server's
`Retry-After`, and API failures surface as typed R conditions you can
`tryCatch` on.

### Installation

From CRAN:

```{r}
install.packages("virustotal")
```

Development version from GitHub:

```{r}
# install.packages("pak")
pak::pak("themains/virustotal")
```

### Quick start

Get a key from your [VirusTotal profile](https://www.virustotal.com/), then:

```{r}
library(virustotal)
set_key("your_api_key")   # or set VIRUSTOTAL_API_KEY in .Renviron

# What does VirusTotal know about a domain, an IP, a file hash, a URL?
domain_report("google.com")
ip_report("8.8.8.8")
file_report("99017f6eebbac24f351415dd410d522d")
url_report("http://www.google.com")

# Submit new things for analysis
scan_url("https://example.org")
scan_file("suspicious.exe")

# Errors are typed conditions
tryCatch(
  file_report("no_such_hash_0000"),
  virustotal_error = function(e) message("VT said: ", conditionMessage(e))
)
```

Public API keys are limited to 4 requests/minute and 500/day; the package
throttles itself to match. Premium keys can raise the pace:

```{r}
options(virustotal.requests_per_minute = 1000)
```

The [vignette](https://themains.github.io/virustotal/articles/using_virustotal.html)
walks through the full surface:

```{r}
vignette("using_virustotal", package = "virustotal")
```

### License

Released under the [MIT License](https://opensource.org/license/mit).
